
\cleardoublepage

\begin{flushright}
	2/15 Dansie Street\\
	Greenslopes, QLD 4120\\
	Tel.\ 0422 655 246\\
	\medskip
	\today
\end{flushright}
\begin{flushleft}
	Prof Michael Bruenig\\
	Head of School\\
	School of Information Technology and Electrical Engineering\\
	The University of Queensland\\
	St Lucia, QLD 4072\\
	\bigskip\bigskip
	Dear Professor Bruenig,
\end{flushleft}

In accordance with the requirements of the degree of Bachelor of Engineering in the division of Software Engineering, I present the following thesis entitled ``\thetitle''. This work was performed under the supervision of Dr Larissa Meinicke \& Prof Ian Hayes.

I declare that the work submitted in this thesis is my own, except as
acknowledged in the text and footnotes, and has not been previously
submitted for a degree at The University of Queensland or any other
institution.
\newline
\begin{flushright}
	Yours sincerely,\\
	\medskip
%	\emph{Author's Signature}
	$ \; $\vspace*{5mm}\\
	\medskip
	Joseph Spearritt.
\end{flushright}

\cleardoublepage

\chapter{Acknowledgements}

I would like to acknowledge the work of my supervisors, Dr Larissa Meinicke and Prof Ian Hayes, in keeping this project on track, and shaping it over our many weekly meetings into something beyond what I could have envisioned it becoming when I started. I would also like to acknowledge Dr Raghavendra K. R. for his advice throughout the project and his role in finding this thesis's ultimate direction -- something which took several iterations and significant discussion. I'd like to thank Aidan Goldthorpe for offering advice and a second opinion to me at many points throughout the year even while working on a thesis project of his own, and finally I'd like to thank my peers, my friends and my family for their support over the lifetime of the project.


%Acknowledge your supervisor, preferably with a few short and specific
%statements about his/her contribution to the content and direction of
%the project.  If you collaborated with another student, acknowledge
%your partner's contribution, including any parts of the thesis of
%which s/he was the principal author or co-author; this information can
%be duplicated in footnotes to the chapters or sections to which your
%partner has contributed.  Briefly describe any assistance that you
%received from technical or administrative staff.  Support of family
%and friends may also be acknowledged, but avoid sentimentality---or
%hide it in the dedication.

\chapter{Abstract}

Mechanisms commonly used to control the confidentiality of information within computer programs frequently fail to capture the potential paths by which information may `leak'. Information Flow security provides a framework for reasoning about the confidentiality properties of a program, and `security typed' programming languages can provide provable guarantees that an application respects the desired confidentiality policy. In this thesis, two such security typed extensions to the Java programming language, Java Information Flow (JIF) and Paragon, are employed to develop three `case study' applications which showcase common real-world security properties. The languages are then evaluated and compared in terms of the confidentiality guarantees they provide, and the burden they place upon the application developer.